More equivalences for Open pNets\thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2018-03-20 2016-Weak-FH-Bisim/WeakBisim.tex
weakbisim running example

notation guards in  SV

@@

  \includegraphics[width=7cm]{XFIG/Transport}
  \hspace{0.5cm}
 \includegraphics[width=4.5cm]{XFIG/TransportSpec}
  \caption{Schema and specification of the protocol }  \label{schema:enable-pnets}
%\end{minipage}
  \vspace{1cm}
%\begin{minipage}{6cm}
  \centerline{\includegraphics[width=11cm]{XFIG/ABPSchema}}
  \caption{Composed pNet with the ABP protocol}  \label{schema:enable-composed}
%\end{minipage}

@@

\begin{example}\emph{An open-transition.}
  \label{OT:enable-composed}
  The \texttt{EnableCompL} pNet of Fig. \ref{schema:enable-composed} has 2 controllers and 2 holes. One of its possible open-transition is:

 \smallskip
 $  OT_2  = \openrule{
                       0 \xrightarrow{\delta}_{C_3} 1 ~~
                            0 \xrightarrow{l}_{C_4} 0  ~~                 
                            \xrightarrow{\delta(x4)}_P ~~
                            \xrightarrow{accept(x4)}_Q 
                      }
  %  {\ostate{00} \xrightarrow{\underline{\delta(x1)}} \ostate{10}}
    {A1_0 \OTarrow{\underline{\delta(x4)}} A1_1}
  $
\end{example}

@@

	\noindent
    $\inferrule
    {k\!\in\! K \\SV_k \!=\! \SV{\alpha_m^{m \in I\uplus J}}{\alpha'}{e_b} \\
    	Leaves(\mylangle \pNet_i^{i\in I'}, \overline{\Sort}, \symb{SV}_k^{k\in 
    	K}\myrangle) \!=\! \pLTS_l^{l\in L} \\    	
    	\forall m\!\!\in\!\! I. 	
    {\left(\begin{array}{l}
	{\pNet_m \models\inferrule*[myfraction=\reddottedrule]
    	{\{s_{i}\xrightarrow{a_{i}}_i s_{i}'\}^{i\in I_m^\prime},
    	\{\xrightarrow{b_{j}}_j\}^{j\in J'_m}, \Pred_m, \Post_m}
    	{\ostate{s_{i}^{i \in L_m}} \OTarrow {v_m}
    		\ostate{(s_i^\prime)^{i\in L_m}}}\lor }\\
		{ \pNet_m 
    	 \models
    	\inferrule*[myfraction=\reddottedrule]
    	{\{s_m \xrightarrow{v_m} s_m'\},\emptyset, \Pred_m, \Post_m}
    	{\ostate{s_m} \OTarrow {v_m}
    		\ostate{s_m'}} \land I'_m=\{m\} \land J'_m=\emptyset}
\end{array}\right)}\\
    	%\land
    	%Leaves(\pNet_m) = \overline{\pLTS}^{L_k})  	
     J' = \biguplus_{m\in I}\!\! J'_m \uplus J 	\\
    	\Pred = \bigwedge_{m\in I}\!\! \Pred_m \land
    	\Pred_{\symb{sv}}(SV_k,v_m^{m\in I},b_j^{j\in J},v)\\ 
    		I' = \biguplus_{m\in I}\!\! I_m'
    	\\\forall i\in	L\backslash I'.\,s'_i=s_i \\
    {\tt fresh}(\alpha_m,\alpha',b_j,v) 
    }
    {\mylangle \pNet_i^{i\in I'}, \overline{\Sort}, \symb{SV}_k^{k\in K}\myrangle
    	\models
    	{\inferrule*[myfraction=\reddottedrule]
    		{\{s_i\xrightarrow{a_i}_i s_i^{\prime}\}^{i\in I^\prime},
    		\{\xrightarrow{b_j}_j\}^{j\in J^\prime}, \Pred, \uplus_{m\in I_k} 
    		\Post_m}
    		{\ostate{s_i^{i\in L}} \OTarrow {v}
    			\ostate{(s_i^\prime)^{i\in L}}}
    	}
    }
    $
	\medskip
%        \TODO{may be explain how $\Pred(SV,a_i^{i\in I_k},b_j^{j\in
%            J_k},v)$ is built ? You mean more than what is written on previous page????}
	%%    \TODO{I have tentatively added the sort constraint on hole actions, that was
	%%not included in the first version... I'm unsure whether this is the best place to
	%%include it, because it may change the decidability conditions on predicates}

@@

\begin{example} \emph{Using the operational rules to compute
    open-transitions}
  In Fig. \ref{usingrules:OT2} we show the deduction tree used to construct and prove the 
  open transition $OT_2$ of \texttt{EnableCompL} (see example page \pageref{OT:enable-composed}).
  The rule uses TR1 for the $\delta$ transition of $C_3$, for the $l$ transition of $C_4$, then combines the result using the $a_4$ vector of the bottom pNet node, and the $\underline{\delta(x)}$ vector of the top node.

@@

2b) For each of the obtained deduction trees, extract the resulting
open-transition, with its predicate and \Post\ assignments by exploring the structure of 
the pNet.

2c)
\ERIC{The predicate is submitted to Z3 for checking satisfiability. If it is NOT stisfiable, the resulting OT is discarded. This will minimize the number of
resulting transitions, and potentialy prune the search-space.}

2d)
\ERIC{Optionally, simplify the predicate by eliminating the unnecessary intermediate variables, that were produced as fresh local variables of synchronisation vectors, and fresh result variables of intermediate OTs. In the resulting predicate the only significant variables are :\\
- the input variables of pLTS transitions\\
- the actions of holes\\
- the result action at toplevel.}

For each open-transition,  add the transition in the outgoing transitions of
the current state, and add the
resulting state in the unexplored set if it is not already in the
explored set.

@@

@@

 Then 
$\mathcal{R}$ is an FH-bisimulation iff for any  states
$s\in\mathcal{S}_1$ and $t\in\mathcal{S}_2$, $(s,t|\Pred_{s,t})\in\mathcal{R}$, we 
have
the following:
\end{minipage}
\hspace{2mm}
\begin{minipage}{0.30\linewidth}
	\includegraphics[width=\linewidth]{XFIG/Bisim}
	\TODO{J'ai ajoute les predicats. Il reste a dessiner la ``couverture'' si j'y arrive...}
\end{minipage}

@@

     \TODO{Eric: j'ai ajoute $\exists$ sur les predicats $\Pred_{s',t_x}$, qui n'etaient pas definis...}

@@

\begin{definition}[Weak open transitions]
  Let $A = <\overline{\LTS},J,\mathcal{S},s_0,\mathcal{T}>$ be an open automaton.
  We define $\mathcal{T^*}$ the \emph{weak open transition relation}
  with respect to a set of invisible actions $\mathcal{I}$ as:

@@

@@

Using the usual encoding of the Lotos Enable operator (see
Fig. \ref{schema:enable-pnets}), we build now open automata for two  
open pNets \texttt{P>>(Q>>R)} and \texttt{(P>>Q)>>R)}, and 
prove their bisimilarity.
\begin{figure}[t]
\centerline{ % \includegraphics[width=0.5\linewidth]{XFIG/Struct-P-QR}
% \includegraphics[width=0.35\linewidth]{XFIG/Struct-PQ-R}}
  \includegraphics[width=0.60\linewidth]{XFIG/P-QR}
  \hspace{5mm}
 \includegraphics[width=0.27\linewidth]{XFIG/PQ-R}}
  \caption{Structure of the pNets for ``P>>(Q>>R) and (P>>Q)>>R''}
  \label{schema:enable-assoc}
\end{figure}

@@

	\noindent
    $\inferrule
    {k\!\in\! K \\SV_k \!=\! \SV{\alpha_m^{m \in I\uplus J}}{\alpha'}{e_b} \\
    	Leaves(\mylangle \pNet_i^{i\in I'}, \overline{\Sort}, \symb{SV}_k^{k\in 
    	K}\myrangle) \!=\! \pLTS_l^{l\in L} \\    	
    	\forall m\!\!\in\!\! I. 	
    {\left(\begin{array}{l}
	{\pNet_m \models\inferrule*[myfraction=\reddottedrule]
    	{\{s_{i}\xrightarrow{a_{i}}_i s_{i}'\}^{i\in I_m^\prime},
    	\{\xrightarrow{b_{j}}_j\}^{j\in J'_m}, \Pred_m, \Post_m}
    	{\ostate{s_{i}^{i \in L_m}} \OTarrow {v_m}
    		\ostate{(s_i^\prime)^{i\in L_m}}}\lor }\\
		{ \pNet_m 
    	 \models
    	\inferrule*[myfraction=\reddottedrule]
    	{\{s_m \xrightarrow{v_m} s_m'\},\emptyset, \Pred_m, \Post_m}
    	{\ostate{s_m} \OTarrow {v_m}
    		\ostate{s_m'}} \land I'_m=\{m\} \land J'_m=\emptyset}
\end{array}\right)}\\
    	%\land
    	%Leaves(\pNet_m) = \overline{\pLTS}^{L_k})  	
     J' = \biguplus_{m\in I}\!\! J'_m \uplus J 	\\
    	\Pred = \bigwedge_{m\in I}\!\! \Pred_m \land
    	\Pred_{\symb{sv}}(SV_k,v_m^{m\in I},b_j^{j\in J},v)\\ 
    		I' = \biguplus_{m\in I}\!\! I_m'
    	\\\forall i\in	L\backslash I'.\,s'_i=s_i \\
    {\tt fresh}(\alpha_m,\alpha',b_j,v) 
    }
    {\mylangle \pNet_i^{i\in I'}, \overline{\Sort}, \symb{SV}_k^{k\in K}\myrangle
    	\models
    	{\inferrule*[myfraction=\reddottedrule]
    		{\{s_i\xrightarrow{a_i}_i s_i^{\prime}\}^{i\in I^\prime},
    		\{\xrightarrow{b_j}_j\}^{j\in J^\prime}, \Pred, \uplus_{m\in I_k} 
    		\Post_m}
    		{\ostate{s_i^{i\in L}} \OTarrow {v}
    			\ostate{(s_i^\prime)^{i\in L}}}
    	}
    }
    $
	\medskip

@@

Now focus on OTs of the subnets. For each $m\in I\cup \{j_0\}$ we have one of the two 
following OT\footnotemark[\thefootnote]:\\[-2ex]
\[
\pNet_m \models\inferrule*[myfraction=\reddottedrule]
    	{\{s_{i}\xrightarrow{a_{i}}_i s_{i}'\}^{i\in I_m^\prime},
    	\{\xrightarrow{b_{j}}_j\}^{j\in J'_m}, \Pred_m, \Post_m}
    	{\ostate{s_{i}^{i \in L_m}} \OTarrow {v_m}
    		\ostate{(s_i^\prime)^{i\in L_m}}}\]
Or
\[{ \pNet_m 
    	 \models
    	\inferrule*[myfraction=\reddottedrule]
    	{\{s_m \xrightarrow{v_m} s_m'\},\emptyset, \Pred_m, \Post_m}
    	{\ostate{s_m} \OTarrow {v_m}
    		\ostate{s_m'}} \land I'_m=\{m\} \land J'_m=\emptyset}\]

@@

This already ensures the second part of the conclusion of the lemma, i.e. the OT for $Q$ 
if we 
choose\footnotemark[\thefootnote]  $v_Q=v_{j_0}$ ($\Pred\,''= \Pred_{j_0}$). Considering 
the OT of $P$ let $I'= 
\biguplus I'_m=I\setminus L_Q$, $J'=\biguplus J'_m\uplus J
\uplus\{j_0\}=J\setminus\Holes(Q)\uplus \{j_0\}$;  the predicate is 
$\Pred\,'=\bigwedge_{m\in I_k}\Pred_m  \land \Pred_{\symb{sv}}(SV_k,v_m^{m\in 
I},b_j^{j\in J\cup \{j_0\}},v)$
where\footnotemark[\thefootnote] $\Pred_{\symb{sv}}(SV_k,v_m^{m\in I},b_j^{j\in J\cup 
\{j_0\}},v)\Leftrightarrow 
\forall i\in I_k.\, \alpha_i=a_i\land \forall j \in J\cup \{j_0\}.\, \alpha_j=b_j \land 
v=\alpha'
\land g$. Modulo renaming of fresh variables, this is identical to the predicate that 
occurs in 
the source open transition except $\alpha_{j_0}=v_{j_0}$ has been replaced by  
$\alpha_{j_0}=b_{j_0}$.
Thus, $\Pred\subst{v_Q\leftarrow b_{j_0}}=(\Pred\,' 
\land \Pred\,'')$. %\TODO{the argument on renaming has mostly been
                 %removed is it still OK? yes.}. 
Finally, Post 
%conditions are easily split\footnote{\TODO{moved. could
%    even remove...?}Note that if the rules are not simplified, no
%  post-condition range over the variables of $P$ and $Q$ at the same
%  time} 
into conditions of the context $P$ and the pNet $Q$ (they are
builts similarly as they only deal with  
leaves): $\Post=\Post\,'\uplus \Post\,''$. This concludes the 
proof as we checked all the premises of the open transition for both $P$ and $Q$.\qed
\footnotetext{Cloning and freshness introduce alpha-conversion at many points 
of the proof; we 
	only 
	give major arguments concerning alpha-conversion to make the proof readable; in 
	general, fresh variables appear in each transition inside
        terms $b_j$, $v$, and 
	$\Pred$.}
\end{proof}
   \end{small}

@@

       Consider each premise of the open transition (constructed by Tr2 in 
Definition~\ref{def:operationalSemantics}). 
We know each premise is true for $P$ and try to prove the equivalent premise for 
$P[Q]_{j_0}$ (using the open transition of $Q$). 
$K$ and the synchronisation vector are unchanged ($j_0$ is now in the set of sub-pNets); 
$SV_k=\SV{\alpha_j^{j\in I\uplus\{j_0\}\uplus 
	J}}{\alpha'}{e_b}$. $\Leaves(P[Q]_{j_0})=\Leaves(P)\uplus \Leaves(Q)$. $I$ and $J$ 
	are the 
    set of leaves and holes of $P$, $I'$ and $J'$ are the sets of moving leaves and holes 
    in the reduction of $P$. All sub-pNets of 
    must
    be 
reduced:\\[-2ex]%\footnote{again we skip arguments related to renaming of fresh 
%variables}:
\[\forall m\!\!\in\!\! I\cup \{j_0\}. 	
    {\left(\begin{array}{l}
	{\pNet_m \models\inferrule*[myfraction=\reddottedrule]
    	{\{s_{i}\xrightarrow{a_{i}}_i s_{i}'\}^{i\in I_m^\prime},
    	\{\xrightarrow{b_{j}}_j\}^{j\in J'_m}, \Pred_m, \Post_m}
    	{\ostate{s_{i}^{i \in L_m}} \OTarrow {v_m}
    		\ostate{(s_i^\prime)^{i\in L_m}}}\lor }\\
		{ \pNet_m 
    	 \models
    	\inferrule*[myfraction=\reddottedrule]
    	{\{s_m \xrightarrow{v_m} s_m'\},\emptyset, \Pred_m, \Post_m}
    	{\ostate{s_m} \OTarrow {v_m}
    		\ostate{s_m'}} \land I'_m=\{m\} \land J'_m=\emptyset}
\end{array}\right)}\]
the sub-pNet at position $j_0$ is the one filled by $Q$ which offers an open transition 
by hypothesis.
Let $I''=\biguplus_{m\in I} I'_m \uplus I_Q = I' \uplus I_Q$ and $J''= \biguplus_{m\in I} 
J_m \uplus J_Q \uplus J = J'\setminus\{j_0\}\uplus J_Q$. The predicate for the open 
transition is: $\Pred\,''=\bigwedge_{m\in I}\Pred_m \land \Pred\,'
\land \Pred(SV_k,a_i^{i\in I_k}\uplus(j_0\mapsto v_Q),b_j^{j\in J},v)$. By definition 
$\Pred(SV_k,a_i^{i\in I}\uplus(j_0\mapsto v_Q),b_j^{j\in J},v)\Leftrightarrow
	\forall i\in I.\, a_i=\alpha_i\land \forall j \in J.\, b_j=\alpha_j \land 
	\alpha'_k=v\land v_Q=\alpha_{j_0}\land e_b$, finally 
	$\Pred\,''=\Pred\subst{b_{j_0}\leftarrow 
	v_Q} \land \Pred\,'$. The post-condition gathers the post-conditions related to all 
	the leaves. Finally, the composed open transition can be
        built. \qed
        \end{proof}
\end{small}

>>>>

X notation guards in  SV

@@

  \includegraphics[width=7cm]{XFIG/Transport}
  \hspace{0.5cm}
 \includegraphics[width=4.5cm]{XFIG/TransportSpec}
  \caption{Schema and specification of the protocol }  \label{schema:ABP-pnets}
%\end{minipage}
  \vspace{1cm}
%\begin{minipage}{6cm}
  \centerline{\includegraphics[width=11cm]{XFIG/ABP3-Schema}}
  \caption{Composed pNet with the ABP protocol}  \label{schema:ABP-composed}
%\end{minipage}

@@

\begin{example}\emph{An open-transition.}
  \label{OT:ABP}
  The \texttt{ABP} pNet of Fig. \ref{schema:ABP-composed} has 2 controllers and 2 holes. One of its possible open-transition is:

 \smallskip
 $  OT_2  = \openrule{
                       1 \xrightarrow{send(m,sb)}_{Sender} 2 ~~
                            \xrightarrow{in(?m,?b)}_{M1}
                      }
  %  {\ostate{00} \xrightarrow{\underline{\delta(x1)}} \ostate{10}}
    { \OTarrow{\ostate{1,0,0,0}\underline{send(m,b)}} \ostate{2,1,0,0}}
  $
\end{example}

@@

	\noindent
    $\inferrule
    {k\!\in\! K \\SV_k \!=\! \SV{\alpha_m^{m \in I\uplus J}}{\alpha'}{e_b} \\
    	Leaves(\mylangle \pNet_i^{i\in I'}, \overline{\Sort}, \symb{SV}_k^{k\in 
    	K}\myrangle) \!=\! \pLTS_l^{l\in L} \\    	
    	\forall m\!\!\in\!\! I. 	
    {\left(\begin{array}{l}
	{\pNet_m \models\inferrule*[myfraction=\reddottedrule]
    	{\{s_{i}\xrightarrow{a_{i}}_i s_{i}'\}^{i\in I_m},
    	\{\xrightarrow{b_{j}}_j\}^{j\in J_m}, \Pred_m, \Post_m}
    	{\ostate{s_{i}^{i \in L_m}} \OTarrow {v_m}
    		\ostate{(s_i^\prime)^{i\in L_m}}}\lor }\\
		{ \pNet_m 
    	 \models
    	\inferrule*[myfraction=\reddottedrule]
    	{\{s_m \xrightarrow{v_m} s_m'\},\emptyset, \Pred_m, \Post_m}
    	{\ostate{s_m} \OTarrow {v_m}
    		\ostate{s_m'}} \land I_m=\{m\} \land J_m=\emptyset}
\end{array}\right)}\\
    	%\land
    	%Leaves(\pNet_m) = \overline{\pLTS}^{L_k})  	
     J' = \biguplus_{m\in I}\!\! J_m \uplus J 	\\
    	\Pred = \bigwedge_{m\in I}\!\! \Pred_m \land
    	\Pred_{\symb{sv}}(SV_k,v_m^{m\in I},b_j^{j\in J},v)\\ 
    		I' = \biguplus_{m\in I}\!\! I_m
    	\\\forall i\in	L\backslash I'.\,s'_i=s_i \\
    {\tt fresh}(\alpha_m,\alpha',b_j,v) 
    }
    {\mylangle \pNet_i^{i\in I'}, \overline{\Sort}, \symb{SV}_k^{k\in K}\myrangle
    	\models
    	{\inferrule*[myfraction=\reddottedrule]
    		{\{s_i\xrightarrow{a_i}_i s_i^{\prime}\}^{i\in I^\prime},
    		\{\xrightarrow{b_j}_j\}^{j\in J^\prime}, \Pred, \uplus_{m\in I_k} 
    		\Post_m}
    		{\ostate{s_i^{i\in L}} \OTarrow {v}
    			\ostate{(s_i^\prime)^{i\in L}}}
    	}
    }
    $
	\medskip
%        \TODO{may be explain how $\Pred(SV,a_i^{i\in I_k},b_j^{j\in
%            J_k},v)$ is built ? You mean more than what is written on previous page????}
	%%    \TODO{I have tentatively added the sort constraint on hole actions, that was
	%%not included in the first version... I'm unsure whether this is the best place to
	%%include it, because it may change the decidability conditions on predicates}

@@

\begin{example} \emph{Using the operational rules to compute
    open-transitions}
  In Fig. \ref{usingrules:OT2} we show the deduction tree used to construct and prove the 
  open transition $OT_2$ of \texttt{EnableCompL} (see example page \pageref{OT:ABP-composed}).
  The rule uses TR1 for the $\delta$ transition of $C_3$, for the $l$ transition of $C_4$, then combines the result using the $a_4$ vector of the bottom pNet node, and the $\underline{\delta(x)}$ vector of the top node.

@@

2b) For each of the obtained deduction trees:\\
- Extract the resulting open-transition, with its predicate and
\Post\ assignments by exploring the structure of the pNet.

- The predicate is submitted to Z3 for checking satisfiability. If it
is NOT stisfiable, the resulting OT is discarded. This will minimize
the number of resulting transitions, and potentialy prune the search-space.

- For each open-transition,  add the transition in the outgoing transitions of
the current state, and add the
resulting state in the unexplored set if it is not already there.

@@

\subsection{Post-processing: Elimination of intermediate variables}

\ERIC{Optionally, simplify the predicate by eliminating the unnecessary intermediate variables, that were produced as fresh local variables of synchronisation vectors, and fresh result variables of intermediate OTs. In the resulting predicate the only significant variables are :\\
- the input variables of pLTS transitions\\
- the actions of holes\\
- the result action at toplevel.}

@@

 Then 
$\mathcal{R}$ is an FH-bisimulation iff for any  states
$s\in\mathcal{S}_1$ and $t\in\mathcal{S}_2$, $(s,t|\Pred_{s,t})\in\mathcal{R}$, we 
have
the following:
\end{minipage}
\hspace{2mm}
\begin{minipage}{0.30\linewidth}
	\includegraphics[width=\linewidth]{XFIG/Bisim}
%	\TODO{J'ai ajoute les predicats. Il reste a dessiner la ``couverture'' si j'y arrive...}
\end{minipage}

@@

%     \TODO{Eric: j'ai ajoute $\exists$ sur les predicats $\Pred_{s',t_x}$, qui n'etaient pas definis...}

\subsection{Post-processing: Elimination of intermediate variables}

\ERIC{Optionally, simplify the predicate by eliminating the unnecessary intermediate variables, that were produced as fresh local variables of synchronisation vectors, and fresh result variables of intermediate OTs. In the resulting predicate the only significant variables are :\\
- the input variables of pLTS transitions\\
- the actions of holes\\
- the result action at toplevel.}

\subsection{Running example}

\TODO{Eric: full result of the tool execution on the running
  example. Full ? certainly not, just a subset...}

Let $\mathcal{R}=\{(s,t|\Pred_{s,t})\}$ be a relation over the set $\mathcal{S}_1$ and 
$\mathcal{S}_2$ constrained by a predicate.
More precisely, for any pair $(s,t)$, there is a 
   single
      $(s,t|\Pred_{s,t})\in\mathcal{R}$  stating that $s$ and $t$ are related 
      if $\Pred_{s,t}$ 
      is 
      true, i.e. the states are related when the variables in $s$ and $t$ verify the 
      predicate $\Pred_{s,t}$.
 FH-bisimulation is defined formally: 
 \begin{definition}[Strong FH-bisimulation]\label{def-FH-bisim}

 Then 
$\mathcal{R}$ is an FH-bisimulation iff for any  states
$s\in\mathcal{S}_1$ and $t\in\mathcal{S}_2$, $(s,t|\Pred_{s,t})\in\mathcal{R}$, we 
have
the following:
\end{minipage}
\hspace{2mm}
\begin{minipage}{0.30\linewidth}
	\includegraphics[width=\linewidth]{XFIG/Bisim}
%	\TODO{J'ai ajoute les predicats. Il reste a dessiner la ``couverture'' si j'y arrive...}
\end{minipage}

%     \TODO{Eric: j'ai ajoute $\exists$ sur les predicats $\Pred_{s',t_x}$, qui n'etaient pas definis...}

@@

\def\InvAct{\mathcal{Inv}}

\begin{definition}[Weak open transitions]
  Let $A = <\overline{\LTS},J,\mathcal{S},s_0,\mathcal{T}>$ be an open automaton.
  We define $\mathcal{T^*}$ the \emph{weak open transition relation}
  with respect to a set of invisible actions $\InvAct$ as:

@@

\subsection{Running example}
- Detail of a $\tau$* sequence, \\
- construction of the predicate\\
- defn / elimination of the intermediaite variables\\
- construction of the weak OT.

\noindent
$\ostate{1000}\OTarrow{m\_rec(m,x), \phi_1 = [x=s_1\_b \land
    m=s_1\_m]}\ostate{2100}$\\
$\qquad\ostate{2100}$ variables: $\psi_1 = \{m_1\_m:=m; m_1\_x:=x; s_2\_b:=s_1\_b\}$;\\
\\
$\ostate{2100}\OTarrow{\tau <m\_send(m_1\_m,m_1\_x), r\_rec(?m',?b)>,
  \phi_2 = [r_0\_b=b \land m'=m_1\_m \land b=m_1\_x]}\ostate{2001}$\\
$\qquad\ostate{2001}$ variables: $\psi_2 = \{r_1\_b:=r_0\_b; r_1\_m:=m;
...\}$;\\
\\
$\ostate{2001}\OTarrow{v,
  \phi_3 = [v=out(r_1\_m)]}\ostate{2002}$\\
$\qquad\ostate{2002}$ variables: $\psi_3 = \{r_2\_b:=r_1\_b\}$;\\

Resulting Weak Transition:

\noindent
$\ostate{1000}\OTWeakarrow{v,
  \phi = \phi_1\land\phi_2\psi_1\land\phi_3\psi_2\psi_1}\ostate{2002}$\\
\\
$\phi=[x=s_1\_b \land m=s_1\_m\\
  \land r_0\_b=b \land m'=m_1\_m \land b=m_1\_x\\
  \land v=out(r_1\_m)]\\
and \\
\psi= \{r_2\_b:=r_0\_b\}$

Note that in $\phi$ the application of assignments
$\psi_1,\psi_2,\psi_3$ have already ``eliminated'' the intermediate
state variables
$m_1\_m, b=m_1\_x, r_1\_m$, but not the intermediate input variables
$m,m',x,b$. But now $\phi$ includes enough information to eliminate them, giving
a final $\phi'= [v=out(s_1\_m)\land r_0\_b=s_1\_b]$

@@

Using the usual encoding of the Lotos Enable operator (see
Fig. \ref{schema:ABP-pnets}), we build now open automata for two  
open pNets \texttt{P>>(Q>>R)} and \texttt{(P>>Q)>>R)}, and 
prove their bisimilarity.
\begin{figure}[t]
\centerline{ % \includegraphics[width=0.5\linewidth]{XFIG/Struct-P-QR}
% \includegraphics[width=0.35\linewidth]{XFIG/Struct-PQ-R}}
  \includegraphics[width=0.60\linewidth]{XFIG/P-QR}
  \hspace{5mm}
 \includegraphics[width=0.27\linewidth]{XFIG/PQ-R}}
  \caption{Structure of the pNets for ``P>>(Q>>R) and (P>>Q)>>R''}
  \label{schema:enable-assoc}
\end{figure}

@@

	\noindent
    $\inferrule
    {k\!\in\! K \\SV_k \!=\! \SV{\alpha_m^{m \in I\uplus J}}{\alpha'}{e_b} \\
    	Leaves(\mylangle \pNet_i^{i\in I'}, \overline{\Sort}, \symb{SV}_k^{k\in 
    	K}\myrangle) \!=\! \pLTS_l^{l\in L} \\    	
    	\forall m\!\!\in\!\! I. 	
    {\left(\begin{array}{l}
	{\pNet_m \models\inferrule*[myfraction=\reddottedrule]
    	{\{s_{i}\xrightarrow{a_{i}}_i s_{i}'\}^{i\in I_m},
    	\{\xrightarrow{b_{j}}_j\}^{j\in J_m}, \Pred_m, \Post_m}
    	{\ostate{s_{i}^{i \in L_m}} \OTarrow {v_m}
    		\ostate{(s_i^\prime)^{i\in L_m}}}\lor }\\
		{ \pNet_m 
    	 \models
    	\inferrule*[myfraction=\reddottedrule]
    	{\{s_m \xrightarrow{v_m} s_m'\},\emptyset, \Pred_m, \Post_m}
    	{\ostate{s_m} \OTarrow {v_m}
    		\ostate{s_m'}} \land I_m=\{m\} \land J_m=\emptyset}
\end{array}\right)}\\
    	%\land
    	%Leaves(\pNet_m) = \overline{\pLTS}^{L_k})  	
     J' = \biguplus_{m\in I}\!\! J_m \uplus J 	\\
    	\Pred = \bigwedge_{m\in I}\!\! \Pred_m \land
    	\Pred_{\symb{sv}}(SV_k,v_m^{m\in I},b_j^{j\in J},v)\\ 
    		I' = \biguplus_{m\in I}\!\! I_m
    	\\\forall i\in	L\backslash I'.\,s'_i=s_i \\
    {\tt fresh}(\alpha_m,\alpha',b_j,v) 
    }
    {\mylangle \pNet_i^{i\in I'}, \overline{\Sort}, \symb{SV}_k^{k\in K}\myrangle
    	\models
    	{\inferrule*[myfraction=\reddottedrule]
    		{\{s_i\xrightarrow{a_i}_i s_i^{\prime}\}^{i\in I^\prime},
    		\{\xrightarrow{b_j}_j\}^{j\in J^\prime}, \Pred, \uplus_{m\in I_k} 
    		\Post_m}
    		{\ostate{s_i^{i\in L}} \OTarrow {v}
    			\ostate{(s_i^\prime)^{i\in L}}}
    	}
    }
    $
	\medskip

@@

Now focus on OTs of the subnets. For each $m\in I\cup \{j_0\}$ we have one of the two 
following OT\footnotemark[\thefootnote]:\\[-2ex]
\[
\pNet_m \models\inferrule*[myfraction=\reddottedrule]
    	{\{s_{i}\xrightarrow{a_{i}}_i s_{i}'\}^{i\in I_m},
    	\{\xrightarrow{b_{j}}_j\}^{j\in J_m}, \Pred_m, \Post_m}
    	{\ostate{s_{i}^{i \in L_m}} \OTarrow {v_m}
    		\ostate{(s_i^\prime)^{i\in L_m}}}\]
Or
\[{ \pNet_m 
    	 \models
    	\inferrule*[myfraction=\reddottedrule]
    	{\{s_m \xrightarrow{v_m} s_m'\},\emptyset, \Pred_m, \Post_m}
    	{\ostate{s_m} \OTarrow {v_m}
    		\ostate{s_m'}} \land I_m=\{m\} \land J_m=\emptyset}\]

@@

This already ensures the second part of the conclusion of the lemma, i.e. the OT for $Q$ 
if we 
choose\footnotemark[\thefootnote]  $v_Q=v_{j_0}$ ($\Pred\,''= \Pred_{j_0}$). Considering 
the OT of $P$ let $I'= 
\biguplus I_m=I\setminus L_Q$, $J'=\biguplus J_m\uplus J
\uplus\{j_0\}=J\setminus\Holes(Q)\uplus \{j_0\}$;  the predicate is 
$\Pred\,'=\bigwedge_{m\in I_k}\Pred_m  \land \Pred_{\symb{sv}}(SV_k,v_m^{m\in 
I},b_j^{j\in J\cup \{j_0\}},v)$
where\footnotemark[\thefootnote] $\Pred_{\symb{sv}}(SV_k,v_m^{m\in I},b_j^{j\in J\cup 
\{j_0\}},v)\Leftrightarrow 
\forall i\in I_k.\, \alpha_i=a_i\land \forall j \in J\cup \{j_0\}.\, \alpha_j=b_j \land 
v=\alpha'
\land g$. Modulo renaming of fresh variables, this is identical to the predicate that 
occurs in 
the source open transition except $\alpha_{j_0}=v_{j_0}$ has been replaced by  
$\alpha_{j_0}=b_{j_0}$.
Thus, $\Pred\subst{v_Q\leftarrow b_{j_0}}=(\Pred\,' 
\land \Pred\,'')$. %\TODO{the argument on renaming has mostly been
                 %removed is it still OK? yes.}. 
Finally, Post 
%conditions are easily split\footnote{\TODO{moved. could
%    even remove...?}Note that if the rules are not simplified, no
%  post-condition range over the variables of $P$ and $Q$ at the same
%  time} 
into conditions of the context $P$ and the pNet $Q$ (they are
builts similarly as they only deal with  
leaves): $\Post=\Post\,'\uplus \Post\,''$. This concludes the 
proof as we checked all the premises of the open transition for both $P$ and $Q$.\qed
\footnotetext{Cloning and freshness introduce alpha-conversion at many points 
of the proof; we 
	only 
	give major arguments concerning alpha-conversion to make the proof readable; in 
	general, fresh variables appear in each transition inside
        terms $b_j$, $v$, and 
	$\Pred$.}
\end{proof}
   \end{small}

@@

       Consider each premise of the open transition (constructed by Tr2 in 
Definition~\ref{def:operationalSemantics}). 
We know each premise is true for $P$ and try to prove the equivalent premise for 
$P[Q]_{j_0}$ (using the open transition of $Q$). 
$K$ and the synchronisation vector are unchanged ($j_0$ is now in the set of sub-pNets); 
$SV_k=\SV{\alpha_j^{j\in I\uplus\{j_0\}\uplus 
	J}}{\alpha'}{e_b}$. $\Leaves(P[Q]_{j_0})=\Leaves(P)\uplus \Leaves(Q)$. $I$ and $J$ 
	are the 
    set of leaves and holes of $P$, $I'$ and $J'$ are the sets of moving leaves and holes 
    in the reduction of $P$. All sub-pNets of 
    must
    be 
reduced:\\[-2ex]%\footnote{again we skip arguments related to renaming of fresh 
%variables}:
\[\forall m\!\!\in\!\! I\cup \{j_0\}. 	
    {\left(\begin{array}{l}
	{\pNet_m \models\inferrule*[myfraction=\reddottedrule]
    	{\{s_{i}\xrightarrow{a_{i}}_i s_{i}'\}^{i\in I_m},
    	\{\xrightarrow{b_{j}}_j\}^{j\in J_m}, \Pred_m, \Post_m}
    	{\ostate{s_{i}^{i \in L_m}} \OTarrow {v_m}
    		\ostate{(s_i^\prime)^{i\in L_m}}}\lor }\\
		{ \pNet_m 
    	 \models
    	\inferrule*[myfraction=\reddottedrule]
    	{\{s_m \xrightarrow{v_m} s_m'\},\emptyset, \Pred_m, \Post_m}
    	{\ostate{s_m} \OTarrow {v_m}
    		\ostate{s_m'}} \land I_m=\{m\} \land J_m=\emptyset}
\end{array}\right)}\]
the sub-pNet at position $j_0$ is the one filled by $Q$ which offers an open transition 
by hypothesis.
Let $I''=\biguplus_{m\in I} I_m \uplus I_Q = I' \uplus I_Q$ and $J''= \biguplus_{m\in I} 
J_m \uplus J_Q \uplus J = J'\setminus\{j_0\}\uplus J_Q$. The predicate for the open 
transition is: $\Pred\,''=\bigwedge_{m\in I}\Pred_m \land \Pred\,'
\land \Pred(SV_k,a_i^{i\in I_k}\uplus(j_0\mapsto v_Q),b_j^{j\in J},v)$. By definition 
$\Pred(SV_k,a_i^{i\in I}\uplus(j_0\mapsto v_Q),b_j^{j\in J},v)\Leftrightarrow
	\forall i\in I.\, a_i=\alpha_i\land \forall j \in J.\, b_j=\alpha_j \land 
	\alpha'_k=v\land v_Q=\alpha_{j_0}\land e_b$, finally 
	$\Pred\,''=\Pred\subst{b_{j_0}\leftarrow 
	v_Q} \land \Pred\,'$. The post-condition gathers the post-conditions related to all 
	the leaves. Finally, the composed open transition can be
        built. \qed
        \end{proof}
\end{small}
